### 2016-07-09 Renormalization.tex
modelfile explanation

\subsection{The Renormalized Matrix Element}
After having renormalized the matrix element for squark production all UV-divergences are removed. This has been achieved by extending the model file, generated by \texttt{SARAH} with counterterm Feynman rules and the above calculated renormalization constants. The derivation of the counterterm Feynman rules has been performed by the author whereas the implementation into the modelfile was mainly done by Philip Dießner, another member of the group. The renormalization constants for masses and fields are calculated by \texttt{FormCalc} functions like \texttt{MassRC} and \texttt{FieldRC}\cite{FormCalcManual}. Their result has been checked with the above calculated renormalization constants.
\begin{lstlisting}
RenConst[dZFu1[gen_]] :=FieldRC[F[3,{gen}]][[1]]
RenConst[dZFd1[gen_]] :=FieldRC[F[4,{gen}]][[1]]

@@

RenConst[dZGG1] := FieldRC[V[5]]
\end{lstlisting}
To keep track of all divergences, the calculation in \texttt{FormCalc} has been performed with the option: \texttt{Dimension} $\to 0$ which is the only opportunity to really work in $D$ dimensions, i.e. the result contains a quantity \texttt{Dminus4} which equals $-2\epsilon$ in DREG. The absolute squared matrix element coming from \texttt{FormCalc} is then given in terms of scalar integrals, like defined in Appendix \ref{sec:Passarino}. To get a notion for the \texttt{FormCalc} output, the following line shows the  off-shell amplitude of the quark self-energy in terms of loop integrals and \texttt{Dminus4}
%\begin{lstlistings}
Amp[{{F[3, {1, Col1}], k[1], 0, {}}} -> {{F[3, {1, Col2}], k[2], 
     0, {}}}][(1/(\[Pi]^2))
  g3^2 ((1/6 + Dminus4/12) (B0i[bb0, Pair[k[1], k[1]], 0, 0] + 
        B0i[bb1, Pair[k[1], k[1]], 0, 0]) - 
     1/6 B0i[bb1, Pair[k[1], k[1]], MassGlu^2, MassSu[1]^2]) Mat[(
SequenceForm["<", "u", 2, "|", 6, ",", 
FormCalc`k[1], "|", "u", 1, ">"]) SUNT[Col1, Col2]] + (1/(\[Pi]^2))
  g3^2 ((1/6 + Dminus4/12) (B0i[bb0, Pair[k[1], k[1]], 0, 0] + 
        B0i[bb1, Pair[k[1], k[1]], 0, 0]) - 
     1/6 B0i[bb1, Pair[k[1], k[1]], MassGlu^2, MassSu[4]^2]) Mat[(
SequenceForm["<", "u", 2, "|", 7, ",", 
FormCalc`k[1], "|", "u", 1, ">"]) SUNT[Col1, Col2]]]
%\end{lstlistings}
To extract infrared divergences, \texttt{LoopTools} has been used. This  tool provides the function \texttt{SetLambda} which can be set to 0, $-1$ or $-2$ to extract the finite part, single pole or double pole of  scalar integrals. This allows for numerically extracting the coefficients of $\epsilon^i$ with $i \in \left\{ 0,-1,-2 \right\}$ of the absolute squared matrix amplitude. The respective poles of $|\mathcal{M}^{\mathrm{V}}_{\mathrm{MRSSM}}(uu \to \tilde{u}_L\tilde{u}_R)|^2$ are shown in figure \ref{fig:MatrixElement}. Note, that the extraction of the single pole does not only involve setting the option \texttt{SetLamda} $\to-1$ and \texttt{Dminus4}$\to 0$, but also setting \texttt{SetLamda}$\to-2$ and \texttt{Dminus4} $\to -2$ and subtract the term obtained by \texttt{SetLamda}$\to-2$ and \texttt{Dminus4} $\to 0$. In doing so, one accounts for the contraction of double poles with terms of $\mathcal{\epsilon}$. The same logic implies to the extraction of finite terms.\\
\\
In addition the tree-level absolute squared matrix amplitude with leading order $\alpha_s$ is compared with its analogue with next-to-leading order $\alpha_s$ and virtual absolute squared matrix amplitude. The latter is  not given analytically. All $|\mathcal{M}|^2$ are color and spin averaged in the final state and summed over in the initial state.
\begin{figure}[!htpb]
\begin{center}
\includegraphics[scale=.5]{figures/MatrixElement_Poles.png}
\caption{Coefficients of $\mathcal{O}(\epsilon^0)$, $\mathcal{O}(\epsilon^{-1})$ and $\mathcal{O}(\epsilon^{-2})$ terms of the absolute squared matrix element of $uu \to \tilde{u}_L\tilde{u}_R$ in the MRSSM as a function of the Mandelstam variable $t$. External helicities and colors are averaged. The parameters are fixed to $m_{\tilde{q}} = \mu = \unit[1000]{GeV}$, $m_{\tilde{g}} = \unit[2000]{GeV}$, $m_{\sigma} = \unit[5000]{GeV}$ and $\sqrt{s} = \unit[6000]{GeV}$ which is the partonic center of mass energy.}
\end{center}
\end{figure}

>>>>

\subsection{The Renormalized Matrix Element}
This section explains how the renormalization of the RSQCD is implemented in the framework of the automated calculation and gives intermediate results in the form of the renormalized matrix element.

\subsubsection{Implementation into the Model File}
Within the automated calculation, the renormlization of the matrix element for squark production has been achieved by extending the model file, generated by \texttt{SARAH} with counterterm Feynman rules and renormalization constants.\\ 
The derivation of the counterterm Feynman rules has been performed by the author whereas the implementation into the modelfile was mainly done by Philip Dießner, another member of the group. To explain the implementation, the quark-squark-gluino vertex is considered, as an example.
\begin{lstlisting}[style=MyMathematica]
C[F[3, {gt1, ct1}], F[15, {ct2}], -S[13, {gt3, ct3}]] == {{
((-I)*g3*Lam[ct2, ct3, ct1]*(Conjugate[ZUL[gt1, 1]]*ZU[gt3, 1] + Conjugate[ZUL[gt1, 2]]*ZU[gt3, 2] + Conjugate[ZUL[gt1, 3]]*ZU[gt3, 3]))/Sqrt[2],
((-I)*(dZFu1[gt1]/2+dZSu1[gt3]/2+dZGlL1/2+dZgs1+dZgs1restore)*g3*Lam[ct2, ct3, ct1]*(Conjugate[ZUL[gt1, 1]]*ZU[gt3, 1] + Conjugate[ZUL[gt1, 2]]*ZU[gt3, 2] + Conjugate[ZUL[gt1, 3]]*ZU[gt3, 3]))/Sqrt[2]
}, {0,0}}
\end{lstlisting}
The names of the particles within the model are given in the \texttt{FeynArts} manual\cite{FeynArtsManual} when  \texttt{MSSMQCD.mod}, the QCD extention of the MSSM model file is explained:\\ 
\mbox{\texttt{F[3,$\left\{ \right.$gt1, ct1$\left.\right\}$]}} is an up-type quark of the \texttt{gt1}'th generation with color \texttt{ct1} and \texttt{F[15,$\left\{ \right.$ct2$\left.\right\}$]} is a gluino with color \texttt{ct2}. However, in comparison to \cite{FeynArtsManual}, there is a slight difference in the definition of squarks. Whereas the MSSM modelfile \texttt{MSSMQCD.mod} has three indices to label squarks, the model file generated by \texttt{SARAH} uses only two: \texttt{-S[13,$\left\{ \right.$ gt3, ct3$\left.\right\}$]} labels an anti up-type squark of mass eigenstate \texttt{gt3} and color \texttt{ct3}. This means more specifically that \texttt{gt3}$\in \left\{ 1,2,3 \right\}$ labels the ``left-handed" squarks with increasing mass and \texttt{gt3}$\in \left\{ 4,5,6 \right\}$ the ``right-handed`'' squarks with increasing mass.\\
The syntax of the model file, also described in \cite{FeynArtsManual}, is like $\left\{\left\{\mathrm{L},\mathrm{L}^{ct}\right\},\left\{\mathrm{R},\mathrm{R}^{ct}\right\}\right\}$, where the shortcuts are explained in table \ref{tab:Modelfille}.
\begin{table}[H]
\begin{center}
\begin{tabular}{c?c}
Shortcut & Meaning\\
\hlinewd{2pt}
L & coupling proportional to the projector $P_L$ \\
L$^\mathrm{ct}$ & counterterm for L \\
R & coupling proportional to the projector $P_R$ \\
R$^\mathrm{ct}$ & counterterm for R \\
\end{tabular}
\caption{Explanation of the vertex structure in the above given mode extract.}\label{tab:Modelfille}
\end{center}
\end{table}
Within the above given extract of the model file, \texttt{Lam} are the Gell-Mann matrices and \texttt{ZU} and \texttt{ZUL} are mixing matrices which are equated to the unit matrix, as there is no squark mixing in the MRSSM and the CKM-matrix is also approximated with the unit matrix.\\
Equipped with that knowledge, one sees that this Feynman rule is the Feynman rule \circled{16b} from Appendix \ref{sec:FeynmanRules}.\\
\\
The renormalization constants for masses and fields are calculated by \texttt{FormCalc} functions like \texttt{MassRC} and \texttt{FieldRC}\cite{FormCalcManual}. 
\begin{lstlisting}[style=MyMathematica]
RenConst[dZFu1[gen_]] :=FieldRC[F[3,{gen}]][[1]]
RenConst[dZFd1[gen_]] :=FieldRC[F[4,{gen}]][[1]]

@@

RenConst[dZGG1] := FieldRC[V[5]]
\end{lstlisting}
The first two lines encode the field renormalization constant of up- and down-type quarks. The function \texttt{FieldRC} actually calculates two renormalization constants when the argument is a fermion field: the left- and right-handed part. But these terms are identical since electroweak interactions are not considered within the modelfile. It therefore suffices to just pick the first element of the list. This is indicated with the addendum $\texttt{[[1]]}$.\\
The next block of the code provides the renormalization constants of up- and down-type squarks, before the ensuing block deals with the gluino. Here, the predefined function \texttt{xZfL1} (\texttt{xZfR1}) is used to extract the field renormalization constant of the left- (right-) handed component of the Dirac-spinor as they couple differently, like already discussed.\\
Finally the gauge coupling renormalization and the supersymmetry restoring counterterm are implemented, before the gluon field is renormalized. The necessity of the parameter \texttt{FiniteGs} will be explained shortly.\\
\\
The parameter \texttt{FiniteGs} has been introduced as these terms should only contribute if \texttt{SetLambda}$\to 0$. In all other choices \texttt{FiniteGs} has to be replaced with zero instead of one. For more details on how the renormalization works within \texttt{FormCalc}, see the manual \cite{FormCalcManual}.\\

After having renormalized the matrix element for squark production all UV-divergences are removed.\\

Their result has been checked with the above calculated renormalization constants.\\
To keep track of all divergences, the calculation in \texttt{FormCalc} has been performed with the option: \texttt{Dimension} $\to 0$ which is the only opportunity to really work in $D$ dimensions, i.e. the result contains a quantity \texttt{Dminus4} which equals $-2\epsilon$ in DREG. The absolute squared matrix element coming from \texttt{FormCalc} is then given in terms of scalar integrals, like defined in Appendix \ref{sec:Passarino}. To get a notion for the \texttt{FormCalc} output, the following line shows the  off-shell amplitude of the quark self-energy in terms of loop integrals and \texttt{Dminus4}
\begin{lstlisting}[style=MyMathematicaOutput]
Amp[{{F[3, {1, Col1}], k[1], 0, {}}} -> {{F[3, {1, Col2}], k[2], 0, {}}}]
[(1/([Pi]^2)) g3^2 ((1/6 + Dminus4/12) (B0i[bb0, Pair[k[1], k[1]], 0, 0] 
+ B0i[bb1, Pair[k[1], k[1]], 0, 0]) - 1/6 B0i[bb1, Pair[k[1], k[1]], MassGlu^2, MassSu[1]^2]) Mat[([<u2|6 k[1]|u1>]) SUNT[Col1, Col2]] 
+ (1/([Pi]^2)) g3^2 ((1/6 + Dminus4/12) (B0i[bb0, Pair[k[1], k[1]], 0, 0] 
+ B0i[bb1, Pair[k[1], k[1]], 0, 0]) - 1/6 B0i[bb1, Pair[k[1], k[1]], MassGlu^2, MassSu[4]^2]) Mat[(<u2|7 k[1]|u1>]) SUNT[Col1, Col2]]]
\end{lstlisting}
To extract infrared divergences, \texttt{LoopTools} has been used. This  tool provides the function \texttt{SetLambda} which can be set to 0, $-1$ or $-2$ to extract the finite part, single pole or double pole of  scalar integrals. This allows for numerically extracting the coefficients of $\epsilon^i$ with $i \in \left\{ 0,-1,-2 \right\}$ of the absolute squared matrix amplitude. The respective poles of $|\mathcal{M}^{\mathrm{V}}_{\mathrm{MRSSM}}(uu \to \tilde{u}_L\tilde{u}_R)|^2$ are shown in figure \ref{fig:MatrixElement}. Note, that the extraction of the single pole does not only involve setting the option \texttt{SetLamda} $\to-1$ and \texttt{Dminus4}$\to 0$, but also setting \texttt{SetLamda}$\to-2$ and \texttt{Dminus4} $\to -2$ and subtract the term obtained by \texttt{SetLamda}$\to-2$ and \texttt{Dminus4} $\to 0$. In doing so, one accounts for the contraction of double poles with terms of $\mathcal{\epsilon}$. The same logic implies to the extraction of finite terms.\\
\\
In addition the tree-level absolute squared matrix amplitude with leading order $\alpha_s$ is compared with its analogue with next-to-leading order $\alpha_s$ and virtual absolute squared matrix amplitude. The latter is  not given analytically. All $|\mathcal{M}|^2$ are color and spin averaged in the final state and summed over in the initial state.
\begin{figure}[!htpb]
\begin{center}
\includegraphics[scale=.5]{figures/MatrixElement_Poles.png}
\caption{Coefficients of $\mathcal{O}(\epsilon^0)$, $\mathcal{O}(\epsilon^{-1})$ and $\mathcal{O}(\epsilon^{-2})$ terms of the absolute squared matrix element of $uu \to \tilde{u}_L\tilde{u}_R$ in the MRSSM as a function of the Mandelstam variable $t$. External helicities and colors are averaged. The parameters are fixed to $m_{\tilde{q}} = \mu = \unit[1000]{GeV}$, $m_{\tilde{g}} = \unit[2000]{GeV}$, $m_{\sigma} = \unit[5000]{GeV}$ and $\sqrt{s} = \unit[6000]{GeV}$ which is the partonic center of mass energy.}
\end{center}
\end{figure}
